---
title: "Vaccine roll out escape mutation frequency project"
output: html_document
---

---
title: "R Notebook"
output: html_notebook
---


Packages

```{r, echo=FALSE }
#install.packages("rjson")
#install.packages("dplyr")
#install.packages("reshape")
#install.packages("rmarkdown")
#install.packages("plotly")
#install.packages("devtools")
#install.packages("hrbrthemes")
#install.packages("patchwork")
#install.packages("nlme")
#install.packages("lubridate")
#install.packages("corrplot")
#install.packages("tidymv")

```


libraries
```{r, echo=FALSE}
library(rjson)
library(jsonlite)
library(dplyr)
require(reshape)
library(ggplot2)
library(plotly)
library(patchwork) 
library(hrbrthemes)
library(mgcv)
library(lubridate)
library(corrplot)
library(tidymv)

```


**Índice**
1. [Mutation data](#id1)
2. [Vaccination data](#id2)
3. [Plots](#id3)
4. [GAMS](#id4)
<div id='id1' />

# MUTATION DATA

```{r}
json_path <- 'S.E484_data.json'
data <- fromJSON(json_path)
data

```

We gather the Json list in a df with the data that interest us and we add a new column to be called country in which it will be able to distinguish from each of the countries that contain the df. 

```{r}
df_cov <- bind_rows(data, .id = 'location')
df_cov
```


### 1. Iterate through every country in our list of countries

#### a. For each country save sublist as dataframe

```{r}
list_df <- list()
i = 1
for (country in names(data)) {
  df <- as.data.frame(data[[country]])
  list_df[[i]] <- df
  i <- i + 1
}
```


```{r}
typeof(data)
```

```{r}
print(list_df[[1]])
```

#### b. Make a column  in dataframe "country" - add the current country

```{r}
list_df <- list()
i = 1
for (country in names(data)) {
  df <- as.data.frame(data[[country]])
  df$country <- country
  list_df[[i]] <- df
  i <- i + 1
}
```

```{r}
print(list_df[[2]])
```

### 2. Merge all of the dataframes together which are in our list


```{r}
merged_df <- do.call("rbind", list_df)
```

```{r}
merged_df %>% 
  filter(country == "Japan")
```


#VACCINATION DATA

```{r}
df_vac <- read.csv("vaccinations.csv")
```


```{r}
names(df_vac)[names(df_vac) == 'date'] <- 'week'
#df_vac = rename(df_vac, c(date="week"))
names (df_vac)
```

```{r}
#df_cov
#df_vac
```

We merged both data frames by the columns Locations and week.


```{r}
library(dplyr)
df <- inner_join(x = df_cov, y = df_vac, by= c("location","week"))
df
```


#PLOTS

1- Mutation


```{r}
ggplot(data = df_cov, mapping = aes(x = cluster_sequences, total_sequences)) +
  geom_point(mapping=aes(color=location))
```
```{r}
plot_ly(data = df_cov, x=~cluster_sequences, y=~total_sequences, color=~location)
```
```{r}
ggplot(data = df_cov, aes(x=week, y=total_sequences))+
geom_line()+
guides(colour=F)+
theme(axis.text.x=element_text(angle = -90, hjust = 0))
```



a.	Plot x = time, y = proportion vaccinated (y = frequency of given mutant)
```{r}
plot_ly(data = df_vac, x=~week, y=~frequency, color=~location)
```


```{r}
prop <- df_vac$people_fully_vaccinated/df_vac$total_vaccinations
plot_ly(data = df_vac, x=~week, y=~people_fully_vaccinated_per_hundred, color=~location)
```




b.	Plotting correlation y=frequency of mutant in given time, x = proportion vaccination in a given month
```{r}
df$prop <- df$cluster_sequences/df$total_sequences
```


```{r}
plot_ly(data = df, y=~prop, x=~people_fully_vaccinated_per_hundred, color=~location)

```

```{r}
plot_ly(data = df, y=~prop, x=~week, color=~location ,type="scatter")
```



1.	Correlation plot (x=proportion vaccination, y=frequency of mutant)


```{r}
frequency <- df$cluster_sequences/df$total_sequences
prop <- df_vac$people_fully_vaccinated/df_vac$total_vaccinations
```


a.	Fit trendlines to the data (similar to covariants.org) - lookup options for doing this in plot_ly or ggplot2 (perhaps the "linetype" parameter for the "plot_ly" function?)

```{r}
plot_ly(data = df, x=~prop, y=~frequency, color=~location) # ,  linetype = ~ 
```


# GAM MODELS

## 1. example with data from Spain

Filter the dataframe to one country (eg UK, US or Spain) so that we can try fitting a GAM to a single country to start with?

```{r}
spain <- df %>% 
  filter(location == "Spain")
```

Add a column to our dataframe for the variant frequency (if this doesn't already exist)

```{r}
frequency <- df$cluster_sequences/df$total_sequences
df <- cbind(df, frequency)
```

 variant freq (y, linear term no. or logisitc transform or fraction of variants) = time + vaccination rate changing over time (predictors)
Model 1: variant freq (y) = time (x)
Model 2: variant freq (y) = time (x) + vaccination rate (x2)


```{r}
Fechas <-as.Date(spain$week, tryFormats = "%Y-%m-%d")
decimal_date(Fechas)
spain$Time <- decimal_date(Fechas)
```


```{r}
spain$frequency <- spain$cluster_sequences/spain$total_sequences
model1 <- gam(frequency~s(Time), data = spain, method = "REML")
```

### Plot 1 (prediction)


```{r}
plot(model1, pages = 1)
```


### Plot 2 (real data)

```{r}
ggplot(spain, aes(x=Time,frequency))+
  geom_point()
```

```{r}

gam.check(model1)
```



Correlation with all variables

-All the variables must be numeric so we will need to hide those variables with characters.

```{r}
df_cor <- select(spain, -location, -week,-iso_code)
round(cor(df_cor, method = "pearson"),
digits= 2)
```



```{r}
spain_cor <- select(spain, -location, -week,-iso_code)
```


```{r}
round(cor(spain_cor),
  digits = 2 # rounded to 2 decimals
)
```

```{r}
M<-cor(spain_cor)
head(round(M,2))
```


```{r}
corrplot(M, type="full")
```

# Combine plots of the GAM model and raw data points (this will help check the fit for the curves which appeared to be shifted downwards UK (0.01) and Spain (0.05)

Check country names for merging datasets
Generate two lists of country names, from the vaccinations and variants dataframes
Remove all countries which are present in both lists
Look at lists to see which countries are present in both dataframes but with different names eg United States vs USA
Rename these countries in one the dataframes so that they're the same
Remerge the dataframes (you should then expect to see more rows in the merged dataframe)

Cambiamos los nombres en los diferentes df para que coincidan (USA,R.checa)

```{r}
df_cov[df_cov$location=="USA",]$location <- "United States"
```

```{r}
df_cov[df_cov$location=="Czech Republic",]$location <- "Czechia"
```


```{r}
df_cov[df_cov$location=="Sint Maarten",]$location <- "Sint Maarten (Dutch part)"
```


```{r}
unique(df_cov$location)
```

```{r}
unique(df_vac$location)
```

We create 2 lists for the mutation data and the vaccination data

```{r}
unique(df_cov$location)
df_cov_location <- unique(df_cov$location)
df_vac_location <- unique(df_vac$location)
df_vac_location
```

Compare both lists

```{r}
final_list <- intersect(df_cov_location, df_vac_location)
final_list
```

```{r}
df_cov_final <- df_cov[is.element(df_cov$location, final_list),]

df_vac_final <- df_vac[is.element(df_vac$location, final_list),]


```

```{r}
unique(df_vac_final$location)
```

create a dataframe to merge both dataframes

```{r}
library(dplyr)
df_final <- inner_join(x = df_cov_final, y = df_vac_final, by= c("location","week"))
df_final
```

```{r}
unique(df_final$location)
```

We show the values that are not present in both lists

```{r}
setdiff(df_cov_location,df_vac_location)

```

```{r}
setdiff(df_vac_location,df_cov_location)
```


```{r}
Fechas <-as.Date(df_final$week, tryFormats = "%Y-%m-%d")
decimal_date(Fechas)
df_final$Time <- lubridate::decimal_date(Fechas)
```

```{r}
frequency <- df_final$cluster_sequences/df$total_sequences
df_final <- cbind(df_final, frequency)
```

```{r}
df_final$frequency <- df_final$cluster_sequences/df_final$total_sequences
```



GAM models <- spain


```{r}
spainm <- gam(frequency~s(Time), data = spain, method = "REML")
summary(spainm)
```

```{r}
model_p <- predict_gam(spainm, length_out = 30, values = list(x0=list(spain$Time)))
model_p
```
queremos ver que datos estan en comun en ambas tablas

```{r}
list(spain$Time)
list(model_p$Time)
```


```{r}
new<- bind_rows(spain,model_p)
```


```{r}
new %>%
  ggplot(aes(Time, fit)) +
  geom_smooth_ci()+
  geom_point(aes(Time,frequency))

```


# Create a function to generate the GAM plot for any country


```{r}
filtrado_df <- function(df1, country) {
df1 <- df1 %>% filter(location == country) 
return(df1)
}
filter_df <- filtrado_df(df_final, 'Spain')
```


```{r}
modelo_df <- function(filter_df) {
model <- gam(frequency~s(Time), data = filter_df, method = "REML")
model_p <- tidymv::predict_gam(model, length_out = 30 , values = list(x0=list(filter_df$Time)))
modelos <- list(model = model, 
                model_p = model_p)
new<- bind_rows(filter_df,model_p) 
return(new)
}

new <- modelo_df(filter_df)

```



```{r}
plot_df <- function(new){
p <- new %>%
  ggplot2::ggplot(ggplot2::aes(Time, fit)) +
  tidymv::geom_smooth_ci()+
  ggplot2::geom_point(ggplot2::aes(Time,frequency))
return(plotly::ggplotly(p))
}

plot <- plot_df(new)

plot

```


```{r}
plot_country <- function(df_final, country){
  filter_df <- filtrado_df(df_final,country)
  new <- modelo_df(filter_df)
  plot <- plot_df(new)
  return(plot)
}

plot_country(df_final,'Czechia')
```

# Plot with all countries

```{r}
countries <- unique(df_final$location)
countries <- countries[1:10]
filter_dfs <- list()
i = 1
for (country in countries) {
  filter_df <- filtrado_df(df_final,country)
  new <- modelo_df(filter_df)
  filter_dfs[[country]] <- as.data.frame(new)
  i <- i + 1
}

actual_final_df <- do.call("rbind", filter_dfs)

p <- actual_final_df %>%
  ggplot2::ggplot(ggplot2::aes(Time, fit, color=location)) +
  tidymv::geom_smooth_ci()+
  ggplot2::geom_point(ggplot2::aes(Time,frequency, color=location))
plotly::ggplotly(p)

```