---
title: "Vaccine roll out escape mutation frequency project"
author: Cindy Catherine Orbegoso Barrantes
date: Feb, 2022
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

Packages

```{r,echo=FALSE,cache=TRUE}
#install.packages("rjson")
#install.packages("dplyr")
#install.packages("reshape")
#install.packages("rmarkdown")
#install.packages("plotly")
#install.packages("devtools")
#install.packages("hrbrthemes")
#install.packages("patchwork")
#install.packages("nlme")
#install.packages("lubridate")
#install.packages("corrplot")
#install.packages("tidymv")
```

libraries

```{r,warning=FALSE,message=FALSE}
library(rjson)
library(jsonlite)
library(dplyr)
require(reshape)
library(ggplot2)
library(plotly)
library(patchwork) 
library(hrbrthemes)
library(mgcv)
library(lubridate)
library(corrplot)
library(tidymv)
```



**Index**
1. [Mutation data](#id1)
2. [Vaccination data](#id2)
3. [Plots](#id3)
4. [GAMS](#id4)


<div id='id1' />
# MUTATION DATA E483K

```{r}
json_path <- 'S.E484_data2.json'
data <- fromJSON(json_path)
head(data[1])
```

We gather the Json list in a df with the data that interest us and we add a new column to be called country in which it will be able to distinguish from each of the countries that contain the df. 

```{r}
df_cov <- bind_rows(data, .id = 'location')
head(df_cov)
```

### 1. Iterate through every country in our list of countries

#### a. For each country save sublist as dataframe

```{r}
list_df <- list()
i = 1
for (country in names(data)) {
  df <- as.data.frame(data[[country]])
  list_df[[i]] <- df
  i <- i + 1
}
```


```{r}
print(list_df[[1]])
```

#### b. Make a column  in dataframe "country" - add the current country

```{r}
list_df <- list()
i = 1
for (country in names(data)) {
  df <- as.data.frame(data[[country]])
  df$country <- country
  list_df[[i]] <- df
  i <- i + 1
}
```

```{r}
print(list_df[[2]])
```

### 2. Merge all of the dataframes together which are in our list


```{r}
merged_df <- do.call("rbind", list_df)
```

```{r}
merged_df %>% 
  filter(country == "Japan")
```

<div id='id2' />
# VACCINATION DATA

```{r}
df_vac <- read.csv("vaccinations.csv")
```


```{r}
names(df_vac)[names(df_vac) == 'date'] <- 'week'
#df_vac = rename(df_vac, c(date="week"))
names (df_vac)
    
```
#df_cov
#df_vac


We merged both data frames by the columns Locations and week.

```{r}
library(dplyr)
df <- inner_join(x = df_cov, y = df_vac, by= c("location","week"))
df
```

<div id='id3' />
# PLOTS

total sequences vs cluster sequences

```{r}
plot_ly(data = df_cov, x=~cluster_sequences, y=~total_sequences, color=~location)
```
total sequences over the time

```{r}
ggplot(data = df_cov, aes(x=week, y=total_sequences))+
geom_line()+
guides(colour=F)+
theme(axis.text.x=element_text(angle = -90, hjust = 0))
```
People fully vaccinated over the time

```{r}
prop <- df_vac$people_fully_vaccinated/df_vac$total_vaccinations
plot_ly(data = df_vac, x=~week, y=~people_fully_vaccinated_per_hundred, color=~location)
```

Plotting correlation y=frequency of mutant in given time, x = proportion vaccination in a given month

```{r}
df$prop <- df$cluster_sequences/df$total_sequences
```

```{r}
plot_ly(data = df, y=~prop, x=~people_fully_vaccinated_per_hundred, color=~location)

```

Proportion over the time by country

```{r}
plot_ly(data = df, y=~prop, x=~week, color=~location ,type="scatter")
```

<div id='id4' />
# GAM MODELS

## 1. example with data from Spain

Filter the dataframe to one country (eg UK, US or Spain) so that we can try fitting a GAM to a single country to start with?

```{r}
spain <- df %>% 
  filter(location == "Spain")
```

Add a column to our dataframe for the variant frequency (if this doesn't already exist)

```{r}
frequency <- df$cluster_sequences/df$total_sequences
df <- cbind(df, frequency)
```

 variant freq (y, linear term no. or logisitc transform or fraction of variants) = time + vaccination rate changing over time (predictors)
Model 1: variant freq (y) = time (x)
Model 2: variant freq (y) = time (x) + vaccination rate (x2)


```{r}
Fechas <-as.Date(spain$week, tryFormats = "%Y-%m-%d")
decimal_date(Fechas)
spain$Time <- decimal_date(Fechas)
```


```{r}
spain$frequency <- spain$cluster_sequences/spain$total_sequences
model1 <- gam(frequency~s(Time), data = spain, method = "REML")
```

### Plot 1 (prediction)


```{r}
plot(model1, pages = 1)
```


### Plot 2 (real data)

```{r}
ggplot(spain, aes(x=Time,frequency))+
  geom_point()
```

```{r}

gam.check(model1)
```


# Combine plots of the GAM model and raw data points (this will help check the fit for the curves which appeared to be shifted downwards UK (0.01) and Spain (0.05)

Check country names for merging datasets
Generate two lists of country names, from the vaccinations and variants dataframes
Remove all countries which are present in both lists
Look at lists to see which countries are present in both dataframes but with different names eg United States vs USA

Remerge the dataframes (you should then expect to see more rows in the merged dataframe)


Rename these countries in one the dataframes so that they're the same (USA,R.checa)

```{r}
df_cov[df_cov$location=="USA",]$location <- "United States"
```

```{r}
df_cov[df_cov$location=="Czech Republic",]$location <- "Czechia"
```

```{r}
df_cov[df_cov$location=="Sint Maarten",]$location <- "Sint Maarten (Dutch part)"
```

```{r}
unique(df_cov$location)
```

```{r}
unique(df_vac$location)
```

We create 2 lists for the mutation data and the vaccination data

```{r}
unique(df_cov$location)
df_cov_location <- unique(df_cov$location)
df_vac_location <- unique(df_vac$location)
df_vac_location
```

Compare both lists

```{r}
final_list <- intersect(df_cov_location, df_vac_location)
final_list
```

```{r}
df_cov_final <- df_cov[is.element(df_cov$location, final_list),]

df_vac_final <- df_vac[is.element(df_vac$location, final_list),]


```

```{r}
unique(df_vac_final$location)
```

create a dataframe to merge both dataframes

```{r}
library(dplyr)
df_final <- inner_join(x = df_cov_final, y = df_vac_final, by= c("location","week"))
df_final
```

```{r}
unique(df_final$location)
```

We show the values that are not present in both lists

```{r}
setdiff(df_cov_location,df_vac_location)

```

```{r}
setdiff(df_vac_location,df_cov_location)
```


```{r}
Fechas <-as.Date(df_final$week, tryFormats = "%Y-%m-%d")
decimal_date(Fechas)
df_final$Time <- lubridate::decimal_date(Fechas)
```

```{r}
frequency <- df_final$cluster_sequences/df$total_sequences
df_final <- cbind(df_final, frequency)
```

```{r}
df_final$frequency <- df_final$cluster_sequences/df_final$total_sequences
```



GAM models <- spain


```{r}
spainm <- gam(frequency~s(Time), data = spain, method = "REML")
summary(spainm)
```

```{r}
model_p <- predict_gam(spainm, length_out = 30, values = list(x0=list(spain$Time)))
model_p
```
queremos ver que datos estan en comun en ambas tablas

```{r}
list(spain$Time)
list(model_p$Time)
```


```{r}
new<- bind_rows(spain,model_p)
```


```{r}
new %>%
  ggplot(aes(Time, fit)) +
  geom_smooth_ci()+
  geom_point(aes(Time,frequency))

```


# Create a function to generate the GAM plot for any country


```{r}
filtrado_df <- function(df1, country) {
df1 <- df1 %>% filter(location == country) 
return(df1)
}
filter_df <- filtrado_df(df_final, 'Spain')
```


```{r}
modelo_df <- function(filter_df, x_col='Time') {
locaion <- filter_df$location[1]
if (x_col == "Time") {
  model <- gam(frequency~s(Time), data = filter_df, method = "REML")
  model_p <- tidymv::predict_gam(model, length_out = 30 , values = list(x0=list(filter_df$Time)))
} else if (x_col == "Prop") {
  model <- gam(frequency~s(people_vaccinated_per_hundred), data = filter_df, method = "REML")
  model_p <- tidymv::predict_gam(model, length_out = 30 , values = list(x0=list(filter_df$people_vaccinated_per_hundred)))}
else if (x_col == "All"){
  model <- gam(frequency~s(Time,k=3)+people_vaccinated_per_hundred, data= filter_df,method= "REML")
  model_p <- tidymv::predict_gam(model, length_out = 30 )
}
modelos <- list(model = model, 
                model_p = model_p)
model_p$location <- locaion 
new<- bind_rows(filter_df,model_p) 
return(new)
} 

new <- modelo_df(filter_df)

```



```{r}
plot_df <- function(new){
p <- new %>%
  ggplot2::ggplot(ggplot2::aes(Time, fit)) +
  tidymv::geom_smooth_ci()+
  ggplot2::geom_point(ggplot2::aes(Time,frequency))
return(plotly::ggplotly(p))
}

plot <- plot_df(new)

plot

```


```{r}
plot_country <- function(df_final, country){
  filter_df <- filtrado_df(df_final,country)
  new <- modelo_df(filter_df)
  plot <- plot_df(new)
  return(plot)
}

plot_country(df_final,'Czechia')
```

# Plot with all countries

```{r}
countries <- unique(df_final$location)
countries <- countries[1:10] #[1:19]
# countries <- c('Spain', 'United Kingdom', 'United States')
filter_dfs <- list()
i = 1
for (country in countries) {
  filter_df <- filtrado_df(df_final,country)
  new <- modelo_df(filter_df)
  filter_dfs[[country]] <- as.data.frame(new)
  i <- i + 1
}

actual_final_df <- do.call("rbind", filter_dfs)
actual_final_df <- actual_final_df %>% tidyr::drop_na(location)

p <- actual_final_df %>%
  ggplot2::ggplot(ggplot2::aes(Time, fit, color=location)) +
  tidymv::geom_smooth_ci()+
  ggplot2::geom_point(ggplot2::aes(Time,frequency, color=location)) +
  ggplot2::theme_minimal() + 
  ggplot2::labs(x = "Time", y = "Frequency of Variant S.E484")
plotly::ggplotly(p) # , width=600, height=400
```



```{r}
countries <- unique(df_final$location)
countries <- countries[1:10] #[1:19]
# countries <- c('Spain', 'United Kingdom', 'United States')
filter_dfs <- list()
i = 1
for (country in countries) {
  filter_df <- filtrado_df(df_final,country)
  new <- modelo_df(filter_df, x_col='All')
  filter_dfs[[country]] <- as.data.frame(new)
  i <- i + 1
}

actual_final_df <- do.call("rbind", filter_dfs)
actual_final_df <- actual_final_df %>% tidyr::drop_na(location)

p <- actual_final_df %>%
  ggplot2::ggplot(ggplot2::aes(people_vaccinated_per_hundred, fit, color=location)) +
  tidymv::geom_smooth_ci()+
  ggplot2::geom_point(ggplot2::aes(people_vaccinated_per_hundred,frequency, color=location)) +
  ggplot2::theme_minimal() + 
  ggplot2::labs(x = "People Vaccinated Per Hundred", y = "Frequency of Variant S.E484")
plotly::ggplotly(p) # , width=600, height=400
```


```{r}

# Single country well
df <- filtrado_df(df_final, 'Spain')
model <- gam(frequency ~ Time + s(people_vaccinated_per_hundred, by = Time), data = df)

# Multiple countries badly
#df <- df_final
#countries <- unique(df_final$location)[1:10]
#df <- subset(df, location %in% countries)
#model <- gam(frequency ~ Time + s(people_vaccinated_per_hundred, by = Time) + location, data = df)

model_p <- tidymv::predict_gam(model)
df <- bind_rows(df, model_p) 

summary(model)

plot(model, shade = TRUE, pages = 1, scale = 0)

df %>%
  ggplot2::ggplot(ggplot2::aes(people_vaccinated_per_hundred, fit, color=location)) +
  ggplot2::geom_point(ggplot2::aes(people_vaccinated_per_hundred,frequency, color=location)) +
  tidymv::geom_smooth_ci()+
  ggplot2::theme_minimal() + 
  ggplot2::labs(x = "People Vaccinated Per Hundred", y = "Frequency of Variant S.E484")
```
Wed 15/12/21



Install R package `gratia`

    
```{r}
install.packages("gratia")
library(gratia)
```


Fit a GAM model for all countries, predicting variant frequency using a tensor product of time and proportion vaccinated using a location (country) as a covariante


# GAM Model 2

```{r}

# df should be a dataframe containing the vaccination + variant data for all countries you want to analyse

df <- df_final
df

countries <- unique(df_final$location)[1:10]
countries

df <- subset(df, location %in% countries)
df

df$location <- as.factor(df$location)


#modelo2 <- gam(frequency ~ location + te(Time, by=people_vaccinated_per_hundred), data = df , family=betar(link='logit'))



modelo2 <- gam(frequency ~ location + te(Time, by=location), data = df, family=betar(link='logit'))


plot(modelo2)
summary(modelo2)


```


Plot the results of the above GAM model using the gratia R package as per the example below

```{r}
library(dplyr)
library(qgam)
library(mgcv)
library(visreg)
#visreg(modelo2, xvar = "frequency", by = "people_vaccinated_per_hundred")
draw(modelo2)
plot(modelo2)
summary(modelo2)
```

24-Jan-2022(should I remove this??)


```{r}
plot_smooth(modelo2, view="Time", plot_all="location",
            rm.ranef=TRUE)
```

example:


```{r}
set.seed(999)
n_obs <- 80
time_span <- 100
time <- sort(runif(n_obs,0,time_span))
Wiener <- cumsum(rnorm(n_obs)) / sqrt(n_obs)
y_obs <- Wiener + rnorm(n_obs,0,.05)

```

```{r}
create.bspline.basis <- function (rangeval=NULL, nbasis=NULL,
                                  norder=4,      breaks=NULL,
                                  dropind=NULL,  quadvals=NULL,
                                  values=NULL,   basisvalues=NULL,
                                  names="bspl")
{
#  This function creates a bspline functional data basis.
#  Arguments
#  RANGEVAL...an array of length 2 containing the lower and upper
#             boundaries for the rangeval of argument values,
#             or a positive number, in which case command
#             rangeval <- c(0, rangeval) is executed.
#             the default is c(0,1)
#  NBASIS  ...the number of basis functions.  This argument must be
#             supplied, and must be a positive integer.
#  NORDER  ...order of b-splines (one higher than their degree).  The
#             default of 4 gives cubic splines.
#  BREAKS  ...also called knots, these are a non-decreasing sequence
#             of junction points between piecewise polynomial segments.
#             They must satisfy BREAKS[1] = RANGEVAL[1] and
#             BREAKS[NBREAKS] = RANGEVAL[2], where NBREAKS is the total
#             number of BREAKS.  There must be at least 2 BREAKS.
#  There is a potential for inconsistency among arguments NBASIS, NORDER,
#  and BREAKS since
#             NBASIS = NORDER + LENGTH(BREAKS) - 2
#  An error message is issued if this is the case.  Although previous
#  versions of this function attempted to resolve this inconsistency in
#  various ways, this is now considered to be too risky.
#  DROPIND ...A vector of integers specifiying the basis functions to
#             be dropped, if any.  For example, if it is required that
#             a function be zero at the left boundary, this is achieved
#             by dropping the first basis function, the only one that
#             is nonzero at that point.
#  QUADVALS...A NQUAD by 2 matrix.  The firs t column contains quadrature
#             points to be used in a fixed point quadrature.  The second
#             contains quadrature weights.  For example, for (Simpson"s
#             rule for (NQUAD = 7, the points are equally spaced and the
#             weights are delta.*[1, 4, 2, 4, 2, 4, 1]/3.  DELTA is the
#             spacing between quadrature points.  The default is
#             matrix("numeric",0,0).
#  VALUES ... A list, with entries containing the values of
#             the basis function derivatives starting with 0 and
#             going up to the highest derivative needed.  The values
#             correspond to quadrature points in QUADVALS and it is
#             up to the user to decide whether or not to multiply
#             the derivative values by the square roots of the
#             quadrature weights so as to make numerical integration
#             a simple matrix multiplication.
#             Values are checked against QUADVALS to ensure the correct
#             number of rows, and against NBASIS to ensure the correct
#             number of columns.
#             The default value of is VALUES is vector("list",0).
#             VALUES contains values of basis functions and derivatives at
#             quadrature points weighted by square root of quadrature weights.
#             These values are only generated as required, and only if slot
#             QUADVALS is not matrix("numeric",0,0).
#  BASISVALUES...A vector of lists, allocated by code such as
#             vector("list",1).
#             This field is designed to avoid evaluation of a
#             basis system repeatedly at a set of argument values.
#             Each list within the vector corresponds to a specific set
#             of argument values, and must have at least two components,
#             which may be tagged as you wish.
#             The first component in an element of the list vector contains the
#             argument values.
#             The second component in an element of the list vector
#             contains a matrix of values of the basis functions evaluated
#             at the arguments in the first component.
#             The third and subsequent components, if present, contain
#             matrices of values their derivatives up to a maximum
#             derivative order.
#             Whenever function getbasismatrix is called, it checks
#             the first list in each row to see, first, if the number of
#             argument values corresponds to the size of the first dimension,
#             and if this test succeeds, checks that all of the argument
#             values match.  This takes time, of course, but is much
#             faster than re-evaluation of the basis system.  Even this
#             time can be avoided by direct retrieval of the desired
#             array.
#             For example, you might set up a vector of argument values
#             called "evalargs" along with a matrix of basis function
#             values for these argument values called "basismat".
#             You might want too use tags like "args" and "values",
#             respectively for these.  You would then assign them
#             to BASISVALUES with code such as
#               basisobj$basisvalues <- vector("list",1)
#               basisobj$basisvalues[[1]] <-
#                               list(args=evalargs, values=basismat)
#  BASISFNNAMES ... Either a character vector of length NABASIS
#             or a single character string to which NORDER, "." and
#             1:NBASIS are appended by the command
#                paste(names, norder, ".", 1:nbreaks, sep="").
#             For example, if norder = 4, this defaults to
#                     'bspl4.1', 'bspl4.2', ... .
#  Returns
#  BASISFD ...a functional data basis object

#  Last modified  11 February 2015 by Jim Ramsay

#  -------------------------------------------------------------------------
#  Default basis for missing arguments:  A B-spline basis over [0,1] of
#    of specified norder with norder basis functions.
#    norder = 1 = one basis function = constant 1
#    norder = 2 = two basis functions = 2 right triangles,
#      one left, the other right.  They are a basis for straight lines
#      over the unit interval, and are equivalent to a monomial basis
#      with two basis functions.  This B-spline system can be
#      explicitly created with the command
#                create.bspline.basis(c(0,1), 2, 2)
#    norder = 3 = three basis functions:  x^2, x-(x-.5)^2, (x-1)^2
#    norder = 4 = default = 4 basis functions
#      = the simplest cubic spline basis
#  -------------------------------------------------------------------------

  type        <- "bspline"

#  if (nargs()==0) {
#    rangeval    <- c(0,1)
#    nbasis      <- 2
#    params      <- NULL
#    dropind     <- NULL
#    quadvals    <- NULL
#    values      <- NULL
#    basisvalues <- NULL
#    basisobj  <- list(type=type, rangeval=rangeval, nbasis=nbasis,
#                  params=params, dropind=dropind,   quadvals=quadvals,
#                  values=values, basisvalues=basisvalues, names=names)
#    oldClass(basisobj) <- "basisfd"
#    return(basisobj)
#  }

#  ------------------------------------------------------------------------
#                     Set up non-default basis
#  ------------------------------------------------------------------------

##
## 1.  check RANGEVAL
##
#  1.1.  First check breaks is either NULL
#        or is numeric with positive length
#  Breaks <- breaks
  op <- options(warn=-1)
  Breaks <- as.numeric(breaks)
  options(op)
  if(!is.null(breaks)){
  	if(min(diff(breaks) < 0)) {
  		stop('One or more breaks differences are negative.')
  	}
    if(is.numeric(breaks)){
      if(length(breaks)<1)breaks <- NULL
      if(any(is.na(breaks)))
        stop('breaks contains NAs;  not allowed.')
      if(any(is.infinite(breaks)))
        stop('breaks contains Infs;  not allowed.')
    }
    else {
#     suppress warning if NAs generated
#      op <- options(warn=-1)
#      Breaks <- as.numeric(breaks)
#      options(op)
      nNA <- sum(is.na(Breaks))
      if(nNA>0)
        stop("as.numeric(breaks) contains ", nNA,
             ' NA', c('', 's')[1+(nNA>1)],
             ';  class(breaks) = ', class(breaks))
    }
  }
#
#  Rangeval <- rangeval
  op <- options(warn=-1)
  Rangeval <- as.numeric(rangeval)
  options(op)
  if(length(rangeval)<1) {
    if(is.null(breaks)) {
      rangeval <- 0:1
    } else{
      rangeval <- range(breaks)
      if(diff(rangeval)==0)
        stop('diff(range(breaks))==0;  not allowed.')
    }
  } else {
#    op <- options(warn=-1)
#    rangeval <- as.numeric(rangeval)
#    options(op)
    nNAr <- sum(is.na(Rangeval))
    if(nNAr>0)
      stop('as.numeric(rangeval) contains ', nNAr,
           ' NA', c('', 's')[1+(nNAr>1)],
           ';  class(rangeval) = ', class(rangeval) )
  }
  if(length(rangeval) == 1){
      if(rangeval <= 0)
        stop("'rangeval' a single value that is not positive, is ",
             rangeval)
      rangeval = c(0,rangeval)
  }
#  if (!rangechk(rangeval)) stop("Argument 'rangeval' is not correct.")
#  if(!is.vector(rangeval))
#    stop('rangeval is not a vector;  class(rangeval) = ',
#         class(rangeval))
# rangeval too long ???
  if(length(rangeval)>2){
    if(!is.null(breaks))
      stop('breaks can not be provided with length(rangeval)>2;  ',
           ' length(rangeval) = ', length(rangeval),
           ' and length(breaks) = ', length(breaks))
    breaks <- rangeval
    rangeval <- range(breaks)
  }
#
  if(rangeval[1]>=rangeval[2])
    stop('rangeval[1] must be less than rangeval[2];  instead ',
         'rangeval[1] = ', rangeval[1], c('==', '>')[diff(rangeval)<0],
         ' rangeval[2] = ', rangeval[2])
##
## 2.  Check norder
##
  if(!is.numeric(norder))
    stop("norder must be numeric;  class(norder) = ",
         class(norder))
#
  if(length(norder)>1)
    stop('norder must be a single number;  length(norder) = ',
         length(norder))
#
  if(norder<=0)stop("norder must be positive, is ", norder)
#
  if((norder%%1) > 0)
    stop("norder must be an integer, = ", norder,
         ', with fractional part = ',norder%%1)
##
## 3.  Check nbasis
##
#  if (is.null(nbasis))     stop("Argument 'nbasis' is not supplied.")
  nbreaks <- length(breaks)
  {
    if(!is.null(nbasis)){
      if(!is.numeric(nbasis))
        stop('nbasis must be numeric, is ', class(nbasis))
      if((lnb <- length(nbasis))>1)
        stop("nbasis must be a single positive integer;  ",
             "length(nbasis) = ", lnb, " > 1;  first 2 elements = ",
             nbasis[1], ", ", nbasis[2])
      if ((nbasis%%1)>0)
        stop("nbasis is not an integer, = ", nbasis,
             ", with fractional part = ", nbasis%%1)
# if (nbasis < 1)          stop("Argument 'nbasis' is not positive.")
      if(nbasis < norder)
        stop('nbasis must be at least norder;  nbasis = ', nbasis,
             ';  norder = ', norder)
##
## 4.  Check breaks
##
#  This argument is optional, and defaults to NULL.
#  if not NULL, it must contain at least two values, the first and last
#  being equal to the corresponding values of RANGEVAL.   The values
#  may not decrease, but there can be sequences of equal values.
#  the number of break values must be consistent with the values
#  of NBASIS and NORDER via the equation
#        NBASIS = NORDER + NBREAKS - 2
      if(!is.null(breaks)){
        if (nbreaks < 2)
          stop("Number of values in argument 'breaks' less than 2.")
        if(breaks[1] != rangeval[1] || breaks[nbreaks] != rangeval[2])
          stop(paste("Range of argument 'breaks' not identical to",
                     "that of argument 'rangeval'."))
        if (min(diff(breaks)) < 0)
          stop("Values in argument 'breaks' are decreasing.")
#  Check for consistency with NBASIS and NORDER
        if (nbasis != norder + nbreaks - 2)
          stop(paste("Relation nbasis = norder + length(breaks) - 2",
                     "does not hold;  nbasis = ", nbasis,
                     "norder = ", norder, "length(breaks) = ",
                     length(breaks)) )
      }
      else{
#  default to nbasis - norder + 2 equally spaced break values
        breaks <- seq(rangeval[1], rangeval[2],
                      length = nbasis - norder + 2)
        nbreaks <- length(breaks)
      }
    }
    else {
#   is.null(nbasis)
      if(is.null(breaks))nbasis <- norder
      else
        nbasis <- length(breaks)+norder-2
    }
  }
##
## 5.  Set up the PARAMS vector, which contains only the interior knots.
##
  if (nbreaks > 2) {
    params <- breaks[2:(nbreaks-1)]
  } else {
    params <- NULL
  }
##
## 6.  set up basis object
##
  basisobj <- basisfd(type=type, rangeval=rangeval, nbasis=nbasis,
                  params=params, dropind=dropind,   quadvals=quadvals,
                  values=values, basisvalues=basisvalues)
##
## 7.  names
##
  {
    ndropind = length(dropind)
    if(length(names) == nbasis)
      basisobj$names <- names
    else {
      if(length(names) > 1)
        stop('length(names) = ', length(names), ';  must be either ',
             '1 or nbasis = ', nbasis)
      basisind = 1:nbasis
      names   = paste(names, norder, ".",as.character(basisind), sep="")
      basisobj$names <- names
    }
  }
##
## 8.  Done
##
##  if(!is.null(axes))basisobj$axes <- axes
  basisobj

}
```




```{r}
times_basis = seq(0,time_span,1)
knots    = c(seq(0,time_span,5)) #Location of knots
n_knots   = length(knots) #Number of knots
n_order   = 4 # order of basis functions: cubic bspline: order = 3 + 1
n_basis   = length(knots) + n_order - 2;

basis = fda::create.bspline.basis(c(min(times_basis),max(times_basis)),n_basis,n_order,knots)
n_basis

```


```{r}
PHI = eval.basis(time, basis) 
dim(PHI)

```

```{r}
matplot(time,PHI,type='l',lwd=1,lty=1, xlab='time',ylab='basis',cex.lab=1,cex.axis=1)
for (i in 1:n_knots)
{
  abline(v=knots[i], lty=2, lwd=1)
}
```

```{r}
# Least squares estimate
# estimate basis coefficient
M = ginv(t(PHI) %*% PHI) %*% t(PHI)
c_hat = M %*% Wiener
```


```{r}
y_hat = PHI %*% c_hat
# Augment data frame for plotting
df <- df %>% mutate(y_hat = y_hat)
p2 <- df %>% ggplot() + 
      geom_line(aes(x = time, y = Wiener), col = "grey") +
      geom_point(aes(x = time, y = y_obs)) +
      geom_line(aes(x = time, y = y_hat), col = "red")
p2 + ggtitle("Original curve and least squares estimate") + 
      xlab("time") + ylab("f(time)")
```

```{r}
# estimate the variance of noise
## SSE = (Y - Xb)'(Y - Xb)
SSE = t(y_hat-y_obs)%*%(y_hat-y_obs)
sigma2 = SSE/(n_obs-n_basis)

# estimate the variance of the fitted curve
# H is the Hat matrix H
# H = X*inv(X'X)*X``
H = PHI %*% M
varYhat = diag(H %*% H * matrix(sigma2,n_obs,n_obs))

# 95% confidence interval

y_hat025 = y_hat-1.96*sqrt(varYhat)
y_hat975 = y_hat+1.96*sqrt(varYhat)
```

```{r}
df <- mutate(df, y_hat025 = y_hat025,
                 y_hat975 = y_hat975)
#names(df) <- c("time","Wiener","y_hat", "y_hat025", "y_hat975")
p3 <- df %>% ggplot() + 
      geom_line(aes(x = time, y = Wiener), col = "grey") +
      geom_point(aes(x = time, y = y_obs)) +
      geom_line(aes(x = time, y = y_hat), col = "red") +
      geom_line(aes(x = time, y_hat025), col = "green") +
      geom_line(aes(x = time, y_hat975), col = "green") 
p3 + ggtitle("Estimated curve with error bars") + 
     xlab("time") + ylab("f(time)")
```


